---
title: 'HW 4'
output: html_document
author: 'Haozhe Chen'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(ARMAacf,dplyr,TinyTeX)
```

1. For each of the following scenarios, choose $\phi_1$ and $\phi_2$
so that the resulting AR(2) process is causal. In each case use R to plot ACF and PACF of the AR(2) process and discuss the characteristics of the plots.
(i) $\phi_1$ > 0 and $\phi_2$ > 0
(ii) $\phi_1$ < 0 and $\phi_2$ > 0
(iii) $\phi_1$ > 0 and $\phi_2$ < 0
(iv) $\phi_1$ < 0 and $\phi_2$ < 0

The condition for causality of AR(2):
$\phi_2+\phi_1<1$, $\phi_2-\phi_1<1$, $-1<\phi_2<1$, so choose the combinations of $\phi_1=(0.5,-0.5)$ and $\phi_2=(0.3,-0.3)$. 
```{r}
# define the function
acfplot = function(phi_1,phi_2){
  y = ARMAacf(ar = c(phi_1,phi_2), lag.max = 20)
  y = y[2:21]
  plot(y, x = 1:20, type = "h", ylim = c(-1,1), xlab = "lags", ylab = "Autocorrelation", main = "AR(2) Population ACF")
abline(h = 0)
  y = ARMAacf(ar = c(phi_1,phi_2), lag.max = 20, pacf=T)
  plot(y, x = 1:20, type = "h", ylim = c(-1,1), xlab = "lags", ylab = "PartialAutocorrelation", main = "AR(2) Population PACF")
abline(h = 0)
}

# plot them
par(mfrow=c(2,2))
acfplot(0.5,0.3)
acfplot(-0.5,0.3)
acfplot(0.5,-0.3)
acfplot(-0.5,-0.3)

# All ACFs decay to zero, either exponentially or like a damped sine wave. 
# All PACFs cut-off after lag 2
```

2.Textbook questions
3.1 Determine which of the following ARMA processes are causal and which of them are invertible. (In each case {Zt} denotes white noise.)
(a) $X_t + 0.2X_{t−1} − 0.48X_{t−2} = Z_t$
$\phi(B)=1+0.2B-0.48B^2$ and $\theta(B)=1$, solve the characteristic equation: 
```{r}
# the roots of phi(B)=0
# the roots lie outside the unit circle
polyroot(c(1,0.2,-0.48))
```
The process is causal is the roots are lie outside the unit circle; The process is invertible becasue the MA part can be expressed as an AR model. 

(b) $X_t + 1.9X_{t−1} + 0.88X_{t−2} = Z_t + 0.2Z_{t−1} + 0.7Z_{t−2}$
$\phi(B)=1+1.9B+0.88B^2$ and $\theta(B)=1+0.2B+0.7B^2$, solve their characteristic equations:
```{r}
# the root of phi(B)=0
polyroot(c(1,1.9,0.88))
# the root of theta(B)=0
polyroot(c(1,0.2,0.7))
# the complex module of theta(B), it is greater than 1
sqrt(0.142857^2+1.186661^2)
```
Hence, the process is not causal, but invertible.

(c) $X_t + 0.6X_{t−1} = Z_t + 1.2Z_{t−1}$
$\phi(B)=1+0.6B$ and $\theta(B)=1+1.2B$, solve their characteristic equations:
```{r}
# the root of phi(B)=0
polyroot(c(1,0.6))
# the root of theta(B)=0
polyroot(c(1,1.2))
```
Since $B_\phi$ are outside the unit circle, $B_\theta$ are inside the unit circle, the process is causal, but not invertible.

(d) $X_t + 1.8X_{t−1} + 0.81X_{t−2} = Z_t$
$\phi(B)=1+1.8B+0.81B^2$ and $\theta(B)=1$
```{r}
# the root of phi(B)=0
polyroot(c(1,1.8,0.81))
```
Thus the process is causal.Also,the process is invertible becasue the MA part can be expressed as an AR model. 

(e) $X_t + 1.6X_{t−1} = Z_t − 0.4Z_{t−1} + 0.04Z_{t−2}$
$\phi(B)=1+1.6B$ and $\theta(B)=1-0.4B+0.04B^2$
```{r}
# the root of phi(B)=0
polyroot(c(1,1.6))
# the root of theta(B)=0
polyroot(c(1,-0.4,0.04))
```
Since $B_\phi$ are inside unit circle, $B_\theta$ are outside the unit circle, the process is not causal, but invertible.

3.2 For those processes in Problem 3.1 that are causal and invertible, compute and graph their ACF and PACF using R.
$$X_t + 0.2X_{t−1} − 0.48X_{t−2} = Z_t$$
```{r}

# ACF and plot
acf = ARMAacf(ar = c(-0.2,0.48), lag.max = 20)
acf = acf[2:21]
plot(acf, x=1:20, type = "h", xlab = "h", ylab = "Autocorrelation", main = "AR(2) Population ACF")
abline(h = 0)

# PACF and plot
pacf = ARMAacf(ar = c(-0.2,0.48), lag.max = 20, pacf=T)
plot(pacf, x=1:20, type = "h", xlab = "h",ylab = "Partial Autocorrelation", main = "AR(2) Population PACF")
abline(h = 0)

```

$$X_t + 1.8X_{t−1} + 0.81X_{t−2} = Z_t$$
```{r}
# ACF and plot
acf = ARMAacf(ar = c(-1.8,-0.81),lag.max = 20)
acf = acf[2:21]
plot(acf, x=1:20, type = "h", xlab = "h", ylab = "Autocorrelation", main = "AR(2) Population ACF")
abline(h = 0)

# PACF and plot
pacf = ARMAacf(ar = c(-1.8,-0.81), lag.max = 20, pacf=T)
plot(pacf, x=1:20, type = "h", xlab = "h",ylab = "Partial Autocorrelation", main = "AR(2) Population PACF")
abline(h = 0)
```

3.4 Compute the ACF and PACF of the AR(2) process $X_t=0.8X_{t−2}+Zt$, ${Z_t}∼WN(0,σ^2)$
This problem was sovled in another document. The R code here is a verification part. 
```{r}
# problem 3.4 --- verify the answers in R 
rho_h = ARMAacf(ar = c(0,0.8), lag.max = 20) 
round(rho_h,3)

phi_hh = ARMAacf(ar = c(0,0.8), pacf=T)
round(phi_hh,3)
```

3. This problem was sovled in another document. The R code here is a verification part. 
```{r}
# problem 3 --- verify the answers in R 
psi_j = ARMAtoMA(ar=c(1.1,-0.3),lag.max = 5)
round(psi_j,3)

rho_h = ARMAacf(ar=c(1.1,-0.3), lag.max = 5)
round(rho_h,3)

phi_hh = ARMAacf(ar=c(1.1,-0.3), pacf=T)
round(phi_hh,3)

```

4. This problem was sovled in another document. The R code here is a verification part. 
```{r}
# problem 4 --- verify the answers in R 
psi_j = ARMAtoMA(ar= 0.6, ma = -0.2, lag.max = 5)
round(psi_j,3)
```

5.This problem was sovled in another document. The R code here is a verification part. 
```{r}
# problem 5 --- verify the answers in R 
rho_h = ARMAacf(ma=c(0.8,-0.15))
round(rho_h,3)

phi_hh = ARMAacf(ma=c(0.8,-0.15),lag.max = 5, pacf=T)
round(phi_hh,3)

```






