---
title: 'HW 6'
output: html_document
author: 'Haozhe Chen'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(TSA)
library(xts)
library(tseries)
```

1. Consider the Johnson and Johnson quarterly earnings data from January 1960 to December 1980 ( R data JohnsonJohnson)
a. Plot the data. Describe the features of the data. Do the data look stationary? Explain your answer.
b. Apply an appropriate variance stabilizing transformation, if necessary.
c. Carry out classical decomposition of the data, plot the transformed series along with the ACF and PACF.
d. Identify an ARMA model for the transformed data.
e. Repeat c and d, but instead of classical decomposition, use differencing to make the data stationary.
```{r}
# JohnsonJohnson
# plot the series
par(mfrow=c(2,2)) 
plot.ts(JohnsonJohnson, main='Johnson and Johnson')
# The series look like a random walk, not stationary. The variance looks at depending on t. 
# apply log transformation
x <- log(JohnsonJohnson)
plot.ts(x, main = 'Johnson and Johnson (log)')

par(mfrow=c(2,2)) 
# apply decomposition
z <- decompose(x, type="additive") 
# plot the stochastic part
plot.ts(z$random, main = 'Johnson and Johnson (Stochastic part)' )
# plot acf and pacf
acf(z$random, na.action = na.pass, lag.max=20, main = 'ACF of Stochastic part')
pacf(z$random, na.action = na.pass, lag.max=20, main = 'PACF of Stochastic part')
# ACF cuts-off after lag 1; PACF cuts-off after lag 3
# The stationary part might be MA(1), AR(3), or ARMA(3,1)
# Apply eacf function
TSA::eacf(na.omit(z$random))
# The stationary part is identified as a MA(1) process

par(mfrow=c(2,2)) 
# Apply differencing, k=1
x1 <- diff(x)
# plot the series 
plot.ts(x1, main = 'Johnson and Johnson (Differencing k=1)')
# plot acf and pacf
acf(x1, na.action = na.pass, lag.max=20, main = 'ACF after Differencing(k=1)')
pacf(x1, na.action = na.pass, lag.max=20, main = 'PACF after Differencing(k=1)')
# ACF cuts-off after lag 1; PACF cuts-off after lag 4
# The stationary part might be MA(1), AR(4), or ARMA(4,1)
# Apply eacf function
TSA::eacf(x1)
# The stationary part is identified as a MA(1) process
```

2. Consider the time series of the numbers of users connected to the Internet through a server every minute (R data WWWusage). Carry out a test for unit root. Apply necessary transformation and identify plausible ARMA models.
```{r}
x <- WWWusage
adf.test(x)
# fail to reject H0, apply transformation

x1 <- log(WWWusage)
adf.test(x1)
# still fail to reject H0, try differencing

# apply differencing to eliminate the trend, k=2
x2 <- diff(diff(WWWusage))
adf.test(x2)
# reject the null, x2 is stationary

# identify the model
par(mfrow=c(2,2)) 
plot.ts(x2)
acf(x2, lag.max=20, main = 'ACF after Differencing (k=2)')
pacf(x2, lag.max=20, main = 'ACF after Differencing (k=2)')
# according to the ACF and PACF plots, x2 might be an ARMA(0,0) process, 
# which means it might be a white noise process

```

6f. Write a R code generate a random sample of size 500 from the AR(1) process with phi = 0.6 and sigma^2 = 0.8, Plot the simulated series along with the sample ACF and PACF of the series. Is the sample ACF and PACF consistent with AR(1)?
```{r}
# simulate the series
AR.sm <- list(order = c(1,0,0), ar = 0.6, sd = sqrt(0.8))
AR1.sm <- arima.sim(n=500, AR.sm)
par(mfrow=c(2,2)) 
plot.ts(AR1.sm, main = 'simulated AR(1)')
acf(AR1.sm, main = 'ACF of simulated series')
pacf(AR1.sm, main = 'PACF of simulated series')
# the ACF dampens to zero and PACF cuts-off at lag 1, 
# so it is consistent with AR(1)
```


# R code for Lecture Notes 4
## Estimation of a Trend: Least Square Approach 
```{r}
par(mfrow=c(2,2)) 
plot.ts(AirPassengers, main="Airline Passenger Data") 
plot.ts(log(AirPassengers), main="Airline Passenger Data after log Transformation") 
# length(AirPassengers)
fit <- lm(log(AirPassengers)~c(1:144)) 
plot(log(AirPassengers), main="Linear Fit to log Transformed Data") 
lines(ts(fit$fitted.values, frequency=12, start=c(1949,1))) 
plot(ts(fit$residuals,frequency=12,start=c(1949,1)),ylab="Residual") 
title(main="Residual of log Airline Pasenger Data") 
```

## Estimation of a Trend: Moving Average Filter 
```{r warning=FALSE}
library(forecast) 
plot(log(AirPassengers), main="Moving Average Smoothing of log AirPassenger Data") 
lines(ma(log(AirPassengers),order=3),col="red") 
lines(ma(log(AirPassengers),order=7),col="blue") 
lines(ma(log(AirPassengers),order=11),col="green") 
legend(1950,6.4,c("3-pt moving ave","7-pt moving average","11-point moving average"),lty=c(1,1), 
col=c("red","blue","green")) 
```

## Estimation of a Trend: Exponential Smoothing 
```{r}
library(forecast) 
fit1 <- ses(log(AirPassengers),alpha=0.6,initial="simple",h=3) 
fit2 <- ses(log(AirPassengers),alpha=0.3,initial="simple",h=3) 
fit3 <- ses(log(AirPassengers),alpha=0.1, initial="simple",h=3) 
plot(log(AirPassengers), main="Exponential Smoothing of log AirPassenger data") 
lines(fitted(fit1),col="red") 
lines(fitted(fit2),col="blue") 
lines(fitted(fit3),col="green") 
legend(1950,6.4,c("lambda=0.6","lambda=0.3","lambda=0.1"), lty=c(1,1),col=c("red","blue","green")) 
```

## Trend via Differencing
```{r}
par(mfrow=c(1,2)) 
plot(log(AirPassengers), main = "log‐AirPassenger Data") 
plot(ts(diff(diff(log(AirPassengers)), lag=12),frequency=12, start=c(1949,1)), 
        ylab="diff(1,12)",main="Differenced Data at lag=(1,12)") 
```

## Additive and Multiplicative Decomposition 
```{r}
z <- decompose(log(AirPassengers), type="additive") 
plot(z) 
# Changing type to “multiplicative” gives multiplicative decomposition. 
# The function decomp.plot allows the title of decompose to be changed. 
decomp.plot <- function(x, main = NULL, ...) 
{ 
  if(is.null(main)) 
    main <- paste("Decomposition of", x$type, "time series") 
    plot(cbind(observed = x$random + 
                 if (x$type == "additive") x$trend + x$seasonal 
               else x$trend * x$seasonal, 
               trend = x$trend, seasonal = x$seasonal, 
    random = x$random), main = main, ...) 
} 
z <- decompose(log(AirPassengers), type="additive") 
decomp.plot(z, main="Additive Decomposition of log(AirPassengers) Data") 
```

## Harmonic Regression Example 
```{r}
library(itsmr) 
x <- trend(log(AirPassengers), 2) 
dlogair <- log(AirPassengers) - x 
y <- hr(dlogair,c(12,6)) 
fit <- x + y 
res <- log(AirPassengers) - fit 
par(mfrow=c(1,2)) 
plot.ts(log(AirPassengers),main="Trend with LM(t , t^2) and HR (d=12 , l=2)") 
lines(ts(fit,frequency=12, star=c(1949,1)),col="red") 
plot.ts(res, ylab="residual", main="Residual") 
```

## Seasonal Differencing 
```{r}
par(mfrow=c(1,2)) 
plot(log(AirPassengers),main="log‐AirPassenger Data") 
plot(ts(diff(diff(log(AirPassengers)),lag=12),frequency=12,start=c(1949,1)), 
        ylab="diff(1,12)",main="Differenced Data at lag=(1,12)") 
```


