---
title: 'HW 7'
output: html_document
author: 'Haozhe Chen'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(dplyr, kableExtra, TSA, forecast)
```


3(a)
For non-seasonal part, ACF decays to zero, PACF cuts-off at lag 1
For seasonal part (which is lag 12), ACF cuts-off at lag 12, PACF (per 12 lags) decays to zero
```{r}
par(mfrow=c(1,2))

plot(ARMAacf(ar = c(0.6), ma = c(0.4, rep(0,10), 0.8),
lag.max=48)[-1],type="h",ylim=c(-0.5,0.5),ylab="ACF",xlab="lag")
abline(h=0)

plot(ARMAacf(ar = c(0.6), ma = c(0.4, rep(0,10), 0.8),
lag.max=48,pacf=T),type="h",ylim=c(-0.5,0.5),
ylab="PACF",xlab="lag")

```

3(b)
ACF cuts-off at lag1 and lag 12, PACF (per 12 lags) decays to zero.  
```{r}
par(mfrow=c(1,2))

plot(ARMAacf( ma = c(0.4, rep(0,10), 0.8, 0.32),
lag.max=48)[-1],type="h",ylim=c(-0.5,0.5),ylab="ACF",xlab="lag")
abline(h=0)

plot(ARMAacf( ma = c(0.4, rep(0,10), 0.8, 0.32),
lag.max=48,pacf=T),type="h",ylim=c(-0.5,0.5),
ylab="PACF",xlab="lag")

```

3(c)
ACF decays to zero either seasonaly and non-seasonaly, the seasonaly lag is 12.
PACF cuts-off at lag1, it becomes significant again around lag 12, and cuts-off at lag 13
```{r}
par(mfrow=c(1,2))

plot(ARMAacf( ar = c(0.4, rep(0,10), 0.8, -0.32),
lag.max=48)[-1],type="h",ylim=c(-0.5,0.5),ylab="ACF",xlab="lag")
abline(h=0)

plot(ARMAacf( ar = c(0.4, rep(0,10), 0.8, -0.32),
lag.max=48,pacf=T),type="h",ylim=c(-0.5,0.5),
ylab="PACF",xlab="lag")

```



4. Consider the co2 data in the dataset pacakage in R, which is Mauna Loa atmospheric CO2 Concentration. Set aside the last 24 observations as the test data and the rest as the training data.
(a) Plot the data and apply Box-Cox transformation, if necessary.
(b) Forecast 1: Use subset selection method to fit an ARIMA model to the data. Verify if the model is adequate.                    Forecast the 24 values along with the forecast intervals.
    Forecast 2: Now identify potential SARIMA models from ACF and PACF plots. Fit the candidate models and
                compare AICC to choose your final model. Use the model to Forecast the 24 values along with the
                forecast intervals.
    Forecast 3: Use Holt-Winters seasonal forecasting method to predict the 24 values along with the forecast                    intervals.
(c) Now complete the following table to compare between the forecasts: 
What is your conclusion?


(a)
```{r}
# co2
plot(co2)
# the variance looks like stable, no need to transform

# head(co2)
```

(b) Forecast 1
```{r}
# partition the data 
train <- co2[1: (length(co2)-24)]
test <- co2[(length(co2)-23): length(co2)]

# the data has monthly seasonalities
fit0 <- armasubsets(y = train, nar = 5, nma = 5)
plot(fit0)
# choose the 5th model

fit1 <- Arima(train, order=c(5,0,3), fixed = c(rep(NA,4),0,rep(NA,4)))
forecast <- forecast(fit1, h = length(test))
plot(forecast)
accuracy(forecast)

```

(b) Forecast 2
```{r}

train_diff1 <- diff(train) # eliminate the trend
train_diff1.12 <- diff(train_diff1, 12) # eliminate the seasonalities

par(mfrow=c(2,2))
acf(train_diff1)
pacf(train_diff1)
acf(train_diff1.12)
pacf(train_diff1.12)

# Comparing following models
# SARIMA(3, 1, 0)x(1, 0, 0)_12
# SARIMA(1, 1, 1)x(0, 1, 1)_12
# SARIMA(1, 1, 0)x(0, 1, 1)_12
# SARIMA(0, 1, 1)x(0, 1, 1)_12

fit2.1 <- Arima(train, order=c(3,1,0), seasonal=list(order=c(1,0,0),period=12), lambda=0)
fit2.2 <- Arima(train, order=c(1,1,1), seasonal=list(order=c(0,1,1),period=12), lambda=0)
fit2.3 <- Arima(train, order=c(1,1,0), seasonal=list(order=c(0,1,1),period=12), lambda=0)
fit2.4 <- Arima(train, order=c(0,1,1), seasonal=list(order=c(0,1,1),period=12), lambda=0)

fit2.1$aicc
fit2.2$aicc
fit2.3$aicc
fit2.4$aicc
# fit2.4 has the lowest AICc, 

fc2 <- forecast(fit2.4, h = 24)
plot(fc2)

# take a look at the accuracy
accuracy(fc2)

```

(b) Forecast 3
```{r}

train_ts <- ts(train, start = c(1959,1), frequency = 12) 
fit3 <- HoltWinters(train_ts, seasonal='additive')
fc3 <- forecast(fit3, 24)
plot(fc3)

# take a look at the accuracy
accuracy(fc3, test)

```


(c)
```{r}

Criteria <- c('Forecast 1','Forecast 2','Forecast 3')
`RMSE (Root Mean Squared Error)` <- c(0.6686277,0.4163387,0.3471728)
`MAPE (Mean Average Percent Error)` <- c(0.1604593,0.07224682,0.07241756)
table <- rbind(Criteria, `RMSE (Root Mean Squared Error)`,`MAPE (Mean Average Percent Error)`)
kable(table,booktabs = T,align = "c")%>%kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

```














